Untitled

### 2017-02-15 Code/calculate_std_mort_rates.Rmd
set 1 agestd rates

```{r load_libraries}
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(reshape2)
library(grid)
library(gridExtra)
```

```{r}
load("~/BW_results/results_set_1.Rdata")
r.All.set1 <- r.All %>% filter(post.samp %in% c(1,2))

for(i in 2:8){
  print(i)
  load(paste0("~/BW_results/results_set_", i, ".Rdata"))
  assign(paste0("r.All.set", i), r.All %>% filter(post.samp %in% c(1,2)))
  rm(r.All)
}

r.All.post1.2 <- rbind(r.All.set1, r.All.set2, r.All.set3, r.All.set4, 
                      r.All.set5, r.All.set6, r.All.set7, r.All.set8)
```

@@

```{r merge}
r.All.post1.2 <- merge(r.All.post1.2, std.pop, by = "age_minbin")

mortality.rates <- r.All.post1.2 %>% 
  group_by(post.samp, state.n, year.n, race.n, sex.n, COD) %>% 
  arrange(age_minbin) %>%
  summarise(age.std.mortality.rate = sum(age.weight*smoothed_rate),
            race = first(race),
            sex = first(sex),
            state = first(state))

ggplot(subset(mortality.rates, post.samp == 1 & state == "Alabama"), aes(y = age.std.mortality.rate, x = year.n)) + 
  geom_line(aes(col = COD)) + facet_grid(race ~ sex)

write.csv(r.All.post1.2, "~/BlackWhiteMortalityGap/Data/states_combined_post_samp_1_2.csv")
write.csv(mortality.rates, "~/BlackWhiteMortalityGap/Data/mortality_rates.csv")
```

>>>>

```{r load_libraries}
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(reshape2)
library(grid)
library(gridExtra)

source("~/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")

```

@@

```{r calculate_age_std_mort_rates_and_bw_diff}
#this take a long time (a few hours!)

system.time(
  for(i in 1:8){
    print(i)
    load(paste0("~/BW_results/results_set_", i, ".Rdata"))
    r.All <- merge(r.All, std.pop, by = "age_minbin")

    #returns dataset with new variable "age.std.mortality.rate" 
    mortality.rates <- r.All %>% 
      group_by(post.samp, state.n, year.n, race.n, sex.n, COD) %>% 
      arrange(age_minbin) %>%
      summarise(age.std.mortality.rate = sum(age.weight*smoothed_rate), #the sum is across age bins
                race = first(race),
                sex = first(sex),
                state = first(state),
                year = first(year))

    mortality.rates.summary <- mortality.rates %>% group_by(state.n, year.n, race.n, sex.n, COD) %>%
      summarise(race = first(race),
                sex = first(sex),
                state = first(state),
                year = first(year),
                age.std.mortality.rate_LCL = quantile(age.std.mortality.rate, 0.025), #summarize across posterior samples (post.samp)
                age.std.mortality.rate_med = quantile(age.std.mortality.rate, 0.5),
                age.std.mortality.rate_UCL = quantile(age.std.mortality.rate, 0.0975))

    #tidy mortality rates summary
    COD.levels <- c("Cardiovascular", "Cancers", "Communicable", 
                    "Non-communicable", "Injuries", "All other causes")

    mortality.rates.summary$COD <- factor(mortality.rates.summary$COD, COD.levels)

    mortality.rates.summary <- add_Census_Division(mortality.rates.summary, state)
    mortality.rates.summary <- add_Census_Region(mortality.rates.summary, state)
    mortality.rates.summary$stabbrs <- state.abb[match(mortality.rates.summary$state, state.name)]
    mortality.rates.summary$rate.per.100k <- mortality.rates.summary$age.std.mortality.rate_med*100000
    mortality.rates.summary <- mortality.rates.summary %>% mutate(state.reorder = reorder(state, as.numeric(Census_Division)))

    #compute black-white difference in mortality
    mortality.rates.white <- mortality.rates %>% filter(race == "White") %>% 
      ungroup() %>%
      select(sex, state, year, COD, post.samp, age.std.mortality.rate) %>%
      rename(age.std.mortality.rate.white = age.std.mortality.rate)

    mortality.rates.black <- mortality.rates %>% filter(race == "Black") %>% 
      rename(age.std.mortality.rate.black = age.std.mortality.rate)

    mortality.rates.wide <- merge(mortality.rates.black, mortality.rates.white, by = c("post.samp", "sex", "state", "year", "COD"))

    mortality.rates.wide <- mortality.rates.wide %>% mutate(rate.difference = age.std.mortality.rate.black - age.std.mortality.rate.white)

    mortality.rates.difference <- mortality.rates.wide %>% 
      group_by(state.n, year.n, sex.n, COD) %>%
      summarise(rate.difference_LCL = quantile(rate.difference, 0.025), #summarize across posterior samples (post.samp)
                rate.difference_med = quantile(rate.difference, 0.5),
                rate.difference_UCL = quantile(rate.difference, 0.0975),
                sex = first(sex),
                state = first(state),
                year = first(year))

    mortality.rates.difference %>% mutate(state.reorder = reorder(state, as.numeric(Census_Division)))

    write.csv(mortality.rates.summary, paste0("~/BlackWhiteMortalityGap/Data/mortality_rates_", i, ".csv"))
    write.csv(mortality.rates.difference, paste0("~/BlackWhiteMortalityGap/Data/mortality_rates_diff_", i, ".csv"))
  }
)
```

### 2017-02-15 Code/plot_age_standardardized_mortality_trends.Rmd
set 1 agestd rates

```{r, echo=F}
mortality.rates <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Data/mortality_rates.csv")
mortality.rates.wide <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Data/mortality_rates_wide.csv")

mortality.rates <- mortality.rates %>% mutate(state.reorder = reorder(state, as.numeric(Census_Division)))
mortality.rates.wide <- mortality.rates.wide %>% mutate(state.reorder = reorder(state, as.numeric(Census_Division)))
```

@@

```{r, echo=F, fig.width=11, eval = F}
for(sex_i in c("Male", "Female")) {
  for(cod in levels(mortality.rates$COD)) {
    for(region in levels(mortality.rates$Census_Region)) {
      data.sub <- subset(mortality.rates, sex == sex_i & post.samp == 1 & COD == cod  & Census_Region == region)
      a <- ggplot(data.sub, 
                      aes(y=rate.per.100k, x = year)) + 
                 geom_vline(aes(xintercept = 1983), col = "#636363", lty = 3) +
                 geom_vline(aes(xintercept = 1993), col = "#636363", lty = 3) +
                 geom_line(aes(col = state, lty = race)) +
                 facet_wrap( ~ state, ncol = 5) +
                 ylab("Age-standardized mortality rate (per 100,000)") +
                 xlab("Year") + 
                 ggtitle(paste0("Age-adjusted mortality (per 100,000) in ", sex_i, " for ", cod, " (Census Region: ", region, ")"))
      print(a)
    }
    for(region in levels(mortality.rates$Census_Region)) {    
          b <- ggplot(subset(mortality.rates.wide, sex == sex_i & post.samp == 1 & COD == cod & Census_Region == region), 
                      aes(y=rate.difference, x = year)) + 
                 geom_vline(aes(xintercept = 1983), col = "#636363", lty = 3) +
                 geom_vline(aes(xintercept = 1993), col = "#636363", lty = 3) +
                 geom_hline(aes(yintercept = 0), col = "#636363") +
                 geom_line(aes(col = state)) +
                 facet_wrap( ~ state, ncol = 5) +
                 ylab("Black-white difference (per 100,000)") +
                 xlab("Year") + 
                 ggtitle(paste0("Black-white difference in age-adjusted mortality (per 100,000) in ", 
                                sex_i, " for ", cod, " (Census Region: ", region, ")"))
          print(b)
    }
  }
}
```

>>>>

```{r, echo=F}
mortality.rates <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Data/mortality_rates.csv")
mortality.rates.wide <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Data/mortality_rates_wide.csv")
```

@@
